---
title: "Common indices"
author: "Mai Lazarus"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---
---
title: "Common indices"
output: html_document
---

This script includes the calculation of commonly used, established indices.

```{r Libraries,echo=T,warning=F,message=F}
library(fractaldim)
library(reshape2)
library(stringr)
library(tidyr)
library(dplyr)
library(features)
```

**Vertical relief (VR) index**

This index represents the difference between the deepest and shallowest measurements in the bottom profile

```{r Vertical relief}
vr_fun = function(depths)  {
  depths_diff = abs(max(depths)-min(depths))
  return(depths_diff)
}
```

**Rugosity index**

This index represents tha ratio between the actual length of the bottom profile contour to the linear, horizontal distance between its start and end points. The length of the contour is calculated by summing the lengths of the hypotenuses of the right angles formed between consecutive measurements.

```{r Rugosity}
rugosity_fun = function(depths,distance)  {
  y_2 = na.omit(depths-lead(depths)) 
  x_dist = distance/length(y_2)
  rugos = sum(sqrt(((x_dist^2)+(y_2^2)))) / distance
  return(rugos)
}
```

**Fractal dimensiond (FDim) index**

The fractal dimension of a physical element is a non-integer dimension that describes its geometric complexity. It represents the change in the bottom profile length with scale, that is, the complexity of the bottom profile shape is expressed as the speed with which its length changes as measurement scale changes (Sugihara & May, 1990). As the fractal dimension increase, the length of the profile changes faster as a function of scale, and so does structural complexity. FDim index was calculated using the Madogram estimator (Gneiting et.al., 2012). 

```{r Fractal dimension}
fractals_fun = function(depths)  {
  data = -1 * depths
  fd = fd.estim.madogram(data, plot.loglog=F, plot.allpoints=F, nlags="auto")
  fractal_d<-as.numeric(fd['fd'])
  return(fractal_d)
}
```

**Slope index**

Slope represents the change in depth as a function of the distance along the bottom profile and is determined by the spatial arrangement of complexity-generating elements. It was calculated as the slope of the linear regression model of depth against distance along the transect.

```{r Slope}
slope_fun = function(depths)  {
  points = c(1:length(depths))
  model = lm(depths~points)
  slope = abs(model$coefficients[2])
  return(slope)
}
```

**Standard deviation index**

The Standard deviation index (SD) is the standard deviation of depths along the bottom profile.

```{r SD}
sd_fun = function(depths)  {
  sd = sd(depths)
  return(sd)
}
```

**Curvature index**

The curvature index describes the mean curvature of a bottom profile, or how strongly the profile changes at the top of convex conFigurations and at the bottom of concave conFigurations (i.e. at the locations of local extrema). It is calculated by fitting a smoothed function to the bottom profile and averaging the curvature values at the local extrema across the transect.

```{r Curvature}
curvature_fun = function(depths)  {
  points = c(1:length(depths))
  feat = features(points, depths, smoother="smooth.spline", spar = 0.05)
  curv = data.frame(c(feat[3]))
  mean_curvature = mean(abs(curv$curvature))
  return(mean_curvature)
}
```


```{r Wrap function,echo=F,eval=T}
common_fun = function (data){
  data = data[,c("ID","depth")]
  common_indices = data %>%
  group_by(ID) %>% 
  summarize(Rugosity = rugosity_fun(depth,distance = 25),
            SD = sd_fun(depth),
            VerRelief = vr_fun(depth),
            Slope = slope_fun(depth),
            FractalDim = fractals_fun(depth),
            Curvature = curvature_fun(depth),
            mean_depth = mean(depth))
  return(common_indices)
}

```

