---
title: "M"
output: pdf_document
---
```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

Install needed packages if not already installed:
```{r Required libraries}

is_inst <- function(pkg) {
  nzchar(system.file(package = pkg))
}

#Install the 'packman' library to easilly install additional libraries
if (!is_inst("pacman")) {install.packages("pacman")}

library("pacman")

p_load(dplyr, tidyverse, knitr ,tidyr, httr, reshape2, stringr, data.table,
       FD, corrplot, fractaldim, features, transport, pracma)

```

This script produces all structural complexity indices used in our study:

1. Structural diversity indices

2. Sky view factor indices

3. Common indices

4. Neighbors distance indices

5. Topographic position indices

6. FreeSpace index

7. EarthMovers index


For each of the groups, an Rmd file is sourced, where all functions are available. A short description of the essence and calculation of each of the aforementioned indices groups is also in the Rmd file. For full description please refer to the manuscript. 

**Data**

The necessary data for all functions should include 3 variables:

  * Depth values along the bottom profile
  * Their order, i.e. an acending sequence of numbers corresponding to the measurements
  * An identification variable distinct for each bottom profile

For functions that require more data explanations are provided.

Step 1: import data 
Here is an example data, including 9000 depth measurements from 50 'transects' 

```{r Load data}
data = read.csv("simulation_data.csv")
head(data)
```

Step 2: calculate structural complexity indices

**1. Structural diversity indices:**

```{r Load Structural diversity functions, echo=FALSE, include=FALSE}
#Load functions
knit('structural diversity functions.Rmd')
```

Calculate structural diversity (FD) indices:
Here, 3 more arguments other than 'data' are required:

1. transect_length - the length of the transect, in m.

2. search distance - the distance in meters in which local extrama are searched from each focal measurement. 

3. traits - a vector with the names of elements traits chosen for the calculation of structural diversity.   

Three types of plots are created, presenting the bottom profile:
For each transect, side by side:

1. Local extrema of all levels. 

2. Structural elements within the profile, distinguished by colors. 

3. A single plot presenting the correlations among structural elements traits. 

*Here, plots 1,2 are shown only for bottom profile number 1. To show the results for all bottom profiles, change the chunks' settings. 

Please refer to 'Structural diversity functions.Rmd' for more details.

```{r Calculate structural diversity, out.width= '50%', results='hide', fig.keep=c(1:2,21)}
FD_indices = indices_calc_fun(data, 
                              transect_length = 25, 
                              search_distance = 0.75, 
                              traits = c("rugosity","vr","max_slope"))
```

```{r}
head(FD_indices)
```


**2. Sky view factor indices**
```{r Load SkyView functions, echo=FALSE, include=FALSE}
#Load functions
knit('Sky view factor functions.Rmd')
```

Calculate Sky view factor (SkyView) indices:
Here, 3 additional arguments are required:
1. 'radius_length' - the length of the arch's radius. 

2. 'threshold' - the amount of angles to extract (%) for the calculation, to avoid the overriding effects of flat areas with low angle values. 

3. 'edges_to_remove' - the distance to remove from each of the transects' ends, as the profile is unknown beyonf the transects' boundaries. 

For full description, refer to 'Sky view factor.Rmd' file and the manuscript. 

```{r Calculate SkyView}
SkyViewFactor = apply_svf_fun(data, radius_length = 2.5, threshold = .3, transect_length = 25, edges_to_remove = 2.5)

head(SkyViewFactor)
```


**3. Common indices**
```{r Load common indices functions, echo=FALSE, include=FALSE}
#Load functions
knit('common_indices_functions.Rmd')
```

Calculate common indices indices:
```{r Calculate common indices}
common_indices = common_fun(data) 
head(common_indices)
```

**4. Neighbors distance indices**

```{r Load Neighbors distance functions, echo=FALSE, include =FALSE}
#Load functions
knit('neighbors_functions.Rmd')
```

Calculate Neighbors distance indices indices:
```{r Calculate Neighbors distance indices}
neighbors_indices = neighbors_indices_fun(data) 
head(neighbors_indices)
```


**5. Topographic position indices - Max concavity/convexity**

```{r Load topographic position indices functions, echo = FALSE, include=FALSE}
#Load functions
knit('TPI functions.Rmd')
```

Calculate Max concavity/convexity indices indices:
```{r Calculate topographical indices}
Max_concave_convex_indices = Max_concave_convex_fun(data)
head(Max_concave_convex_indices)
```

**6. FreeSpace index**

```{r Load FreeSpace functions, echo=FALSE, include=FALSE}
#Load functions
knit('FreeSpace functions.Rmd')
```

Calculate FreeSpace index:
```{r Calculate FreeSpace index}
FreeSpace_index = FreeSpace_fun(data) 
head(FreeSpace_index)
```


**7. EarthMovers index**

```{r Load EarthMovers functions, echo=FALSE, include=FALSE}
#Load functions
knit('EarthMovers functions.Rmd')
```

Calculate EarthMovers indices:
```{r Calculate EarthMovers functions}
EarthMovers_data = wass_fun(data)
head(EarthMovers_data)
```

Join all indices data:
```{r Join all indices}
complexity_indices = Reduce(function(x, y) merge(x, y, by = "ID", all = TRUE), list(FD_indices, SkyViewFactor, common_indices, neighbors_indices, Max_concave_convex_indices, FreeSpace_index, EarthMovers_data))
head(complexity_indices)
```


